---
title: "Prosper Loan Data Analysis"
output: html_notebook
---

## Loading the packages
```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(ggthemes))
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
suppressMessages(library(memisc))
suppressMessages(library(gridExtra))
suppressMessages(library(RColorBrewer))
```

## Loading the dataset
```{r}
loanData <- read.csv('prosperLoanData.csv')
```

## The first Qustion
The first question that needs to be asked is HOW LONG PEOPLE USUALLY OPT FOR LOAN? Let's answer this question with a histogram
```{r}
loanData %>% 
    ggplot(aes(x = Term / 12)) +
        geom_histogram(binwidth = 1) +
        theme_pander() + 
        xlab('Term in Years') +
        ylab('How many?') + 
        scale_x_continuous(breaks = seq(1, 5, 1))
```
We can see that people don't really loan any amount for less than one year and the most popular loan amount is of 3 years although some people do choose for 5 years. Now lets assume something and check if it is correct or not. I assume that people who opted for 1 year would fail to repay their loan more as compared to the people who opeted for 3 or 5 years. But is it true? So the next question is
DOES PEOPLE REPAY THE LOANS BETTER WHEN THEY ARE GIVEN MORE TIME?






